---
title: "R Notebook"
output: html_notebook
---

# Installation of required packages

```{r, eval = FALSE}
install.packages("CausalImpact")
install.packages("tidyverse")
#devtools::install_github("teunbrand/ggh4x")
install.packages("tidyquant")
install.packages("gtrendsR")
```

# load required packages

```{r}
library(tidyverse)
library(lubridate)
library(CausalImpact)
library(janitor)
library(broom)
library(xtable)
library(tidyquant)
library(gtrendsR)
```

Read the monthly data

```{r}
df <- read_csv("../data/gig-monthly-stats-new.csv") %>%
  mutate(month = row_number() %% 12,
         month = ifelse(month == 0, 12, month),
         year = str_extract(app, "\\d{4}"),
         month = paste(15, month, year, sep = "-"),
         month = as.Date(month, "%d-%m-%Y"))  %>%
  dplyr::select(-app, -year) %>%
  gather(apps, vals, -month) %>%
  filter(!apps %in% c("Google Chrome", "Bolt Courier")) %>%
  mutate(apps = case_when(apps == "Bolt Driver" ~ "Bolt",
                          apps == "Glover" ~ "Glover",
                          apps == "DeliGoo Courier" ~ "DeliGoo",
                          apps == "Takeaway.com Courier" ~ "Takeaway",
                          apps == "Uber Driver" ~ "Uber",
                          apps == "Wolt Courier Partner" ~ "Wolt"),
         category = ifelse(apps %in% c("Glover", "Wolt", "Takeaway", "DeliGoo"), "Delivery", "Transport"),
         apps = factor(apps, c("DeliGoo", "Glover", "Takeaway", "Wolt", "Bolt", "Uber"))) %>%
  group_by(apps) %>%
  mutate(trend = row_number() - 27) %>%
  ungroup() %>%
  filter(vals > 0)
  

df
```

Add trend line before pandemic

```{r}
df %>%
  filter(trend  < 0) %>%
  group_by(apps) %>%
  do(model = tidy(lm(vals ~ trend, data = .))) %>%
  unnest(cols = c(model)) %>%
  dplyr::select(apps, term, estimate) %>%
  spread(term, estimate) -> trend_lines

trend_lines
```


```{r}
ggplot(data = df, aes(x = trend, y = vals, group = apps)) +
  geom_line() + 
  geom_abline(data = trend_lines, aes(slope = trend, intercept = `(Intercept)`, group = apps), 
              linetype = "dashed", color = "red") +
  facet_wrap(~  apps, ncol = 3, scales = "free_y") +
  geom_vline(xintercept = 0, linetype = "dotted", color = "black") + 
  geom_vline(xintercept = 6, linetype = "dotted", color = "black") +
  geom_vline(xintercept = 10, linetype = "dotted", color = "black") +
  theme_bw()  +
  scale_x_continuous(breaks = c(-24, -18, -12, -6, 0, 6, 12, 18, 24)) + 
  labs(x = "Months (March 2020=0)", y = "Number of active users",
       caption = "Note: dotted horizontal lines denote begning of a given COVID-19 wave in Poland. The first is March 2020, the second is September 2020\nand the third is January 2021. Red dashed line denote pre-COVID-19 trend (before March 2020)") -> p1
p1
```

```{r}
ggsave(plot = p1, filename = "../figures/fig1-general.pdf", width = 9, height = 5)
```


# Causal Inference using space-state models

We use bayesian space state models to estimate casual impact of the first COVID wave

```{r}
df %>% count(apps) %>% 
  filter(apps != "Takeaway") %>% pull(apps) %>% as.character() -> apps

apps
```

```{r}
space_state_results <- list()

for (a in apps) {
  analysis <- df %>%
    filter(apps %in% c("Takeaway", a)) %>%
    dplyr::select(number_of_users = vals, apps, month) %>%
    filter(number_of_users > 0) %>%
    spread(apps, number_of_users) %>%
    na.omit()  %>%
    mutate(trend = row_number())
  
  analysis_ts <- zoo(cbind(analysis[[a]], analysis$Takeaway), as.Date(analysis$month))
  pre.period <- c(min(analysis$month), as.Date("2020-02-15"))
  post.period <- c(as.Date("2020-03-15"), max(analysis$month))
  space_state_results[[a]] <- CausalImpact(analysis_ts, pre.period, post.period, 
                                           model.args = list(niter = 5000))


}

```

# Summary

```{r}
lapply(space_state_results, plot)
```
Columns AbsEffect, AbsEffect.lower, AbsEffect.upper, RelEffect, RelEffect.lower,  RelEffect.upper

```{r}
space_state_results %>%
  map("series") %>%
  map_df(~.x %>% as.data.frame() %>%
           rownames_to_column(var = "month") %>%
           dplyr::select(month, response, point.effect, point.effect.lower, point.effect.upper) %>%
           filter(month == "2020-08-15"), .id = "apps") %>%
  mutate(rel_m = point.effect/response*100,
         rel_m_l = point.effect.lower/response*100,
         rel_m_u = point.effect.upper/response*100,
         abs_m = as.integer(round(point.effect)),
         abs_m_l = as.integer(round(point.effect.lower)),
         abs_m_u = as.integer(round(point.effect.upper))) %>%
  dplyr::select(apps, abs_m:abs_m_u, rel_m:rel_m_u) -> before_2_wave
  
```

```{r}
before_2_wave %>%
  xtable(digits = 1) %>%
  print.xtable(include.rownames = F, format.args=list(big.mark = ","))
```

## Appendix

### COVID-19 in Poland

```{r}
covid_poland <- read_csv("../data/covid-in-poland.csv")  

ggplot(data = covid_poland, aes(x = date, y = new_cases, group = 1)) +
  geom_line() +
  theme_bw() +
  labs(x = "Date", y = "New cases") +
  geom_vline(xintercept = as.Date("2020-03-16"), linetype = "dashed", color = "black") + 
  geom_vline(xintercept = as.Date("2020-09-30"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2021-01-31"), linetype = "dashed", color = "black") +
  geom_ma(ma_fun = SMA, n = 7, col = "red", linetype = "solid", size = 1) -> p_cases

p_cases
```

```{r}
ggsave(plot = p_cases, filename = "../figures/fig-appen-cases.pdf", width = 9, height = 5)
```

### Google trends

```{r}
googletrends <- gtrends(keyword = c("Na wynos", "Taksówka"), geo = "PL")

googletrends$interest_over_time %>%
  mutate(keyword = ifelse(keyword == "Na wynos", 
                          "food to takeaway\n(pol. na wynos)", "taxi\n(pol. taksówka)")) %>%
  filter(year(date) %in% 2018:2021) %>%
  ggplot(data = ., aes(x = date, y = hits, color = keyword, group = keyword)) +
  geom_line() +
  theme_bw() +
  labs(x= "Date", y = "Index", color = "Keyword") +
  scale_color_brewer(type = "qual", palette = "Set1") -> p1

p1
```
```{r}
ggsave(plot = p1, filename = "../figures/fig-appen-gtrends.pdf", width = 9, height = 5)
```

```{r}
googletrends_apps <- gtrends(keyword = c("Pyszne.pl", "Glovo", "DeliGoo", "Wolt"), geo = "PL")

googletrends_apps$interest_over_time %>%
  #mutate(keyword = ifelse(keyword == "Na wynos", "food to takeaway", "taxi")) %>%
  mutate(hits = as.numeric(hits),
         keyword = factor(keyword, 
                          c("DeliGoo", "Glovo", "Pyszne.pl", "Wolt"),
                          c("DeliGoo", "Glovo", "Takeaway", "Wolt"))) %>%
  filter(year(date) %in% 2018:2021) %>%
  ggplot(data = ., aes(x = date, y = hits, color = keyword, group = keyword)) +
  geom_line() +
  theme_bw() +
  labs(x= "Date", y = "Index", color = "Keyword") +
  scale_color_brewer(type = "qual", palette = "Set1") -> p2

p2
ggsave(plot = p2, filename = "../figures/fig-appen-gtrends-apps.pdf", width = 9, height = 5)
```

```{r}
googletrends_apps <- gtrends(keyword = c("Uber", "Uber Eats"), geo = "PL")

googletrends_apps$interest_over_time %>%
  #mutate(keyword = ifelse(keyword == "Na wynos", "food to takeaway", "taxi")) %>%
  mutate(hits = as.numeric(hits),) %>%
  filter(year(date) %in% 2018:2021) %>%
  ggplot(data = ., aes(x = date, y = hits, color = keyword, group = keyword)) +
  geom_line() +
  theme_bw() +
  labs(x= "Date", y = "Index", color = "Keyword") +
  scale_color_brewer(type = "qual", palette = "Set1") -> p3

p3
ggsave(plot = p3, filename = "../figures/fig-appen-gtrends-apps2.pdf", width = 9, height = 5)
```


### Demographics

Shares between II half of 2019 and II half of 2020.

```{r}
gig_demo <- read_csv("../data/gig-table2-halfyear-demo-stats.csv") %>%
  filter(period %in% c(2019.2, 2020.2)) %>%
  filter(!apps %in% c("Bolt Courier", "FREE NOW", "iTaxi")) %>%
  dplyr::select(gender:n_unique_users) %>%
  mutate(period = ifelse(period == "2019.2", "2019 HY2", "2020 HY2"),
         apps = ifelse(apps == "Bolt Driver", "Bolt", apps))

gig_demo %>%
  count(apps, gender, period, wt = n_unique_users) %>%
  add_count(apps, period, wt = n, name = "total") %>%
  mutate(p = n/total*100) %>%
  dplyr::select(-n,-total) %>%
  spread(apps, p) %>%
  rename(var = gender) -> gig_gender

gig_demo %>%
  count(apps, age, period, wt = n_unique_users) %>%
  add_count(apps, period, wt = n, name = "total") %>%
  mutate(p = n/total*100) %>%
  dplyr::select(-n,-total) %>%
  spread(apps, p, fill = 0) %>%
  rename(var = age) -> gig_age

gig_demo %>%
  count(apps, country, period, wt = n_unique_users) %>%
  add_count(apps, period, wt = n, name = "total") %>%
  mutate(p = n/total*100) %>%
  dplyr::select(-n,-total) %>%
  spread(apps, p, fill = 0) %>%
  rename(var = country) -> gig_country
```


Spatial

```{r}
#gig_spatial <- 
read_csv("../data/gig-table3-halfyear-region-stats.csv") %>%
  filter(period %in% c(2019.2, 2020.2)) %>%
  filter(!apps %in% c("Bolt Courier", "FREE NOW", "iTaxi")) %>%
  count(group, period, apps, wt = number_of_users) %>%
  mutate(period = ifelse(period == "2019.2", "2019 HY2", "2020 HY2"),
         apps = ifelse(apps == "Bolt Driver", "Bolt", apps)) %>%
  spread(group, n) %>%
  mutate(Voivodeship = Voivodeship - `Functional area`,
         `Functional area` = `Functional area` - Cities) %>%
  gather(geo, vals, -period, -apps) %>%
  add_count(period, apps, wt = vals, name = "total") %>%
  mutate(p = vals/total*100) %>%
  dplyr::select(-vals, -total) %>%
  spread(apps, p) %>%
  arrange(geo, period) %>%
  rename(var = geo) -> gig_spatial
```

```{r}
gig_gender %>%
  bind_rows(gig_age) %>%
  bind_rows(gig_country) %>%
  bind_rows(gig_spatial) %>%
  xtable(digits = 1) %>%
  print.xtable(include.rownames = F)
```


## Model results

```{r}
for (a in apps) {
  p <- plot(space_state_results[[a]], c("original", "pointwise"))
  ggsave(plot = p, filename = paste0("../figures/fig-est-", a, ".pdf"), width = 9, height = 5)
}

```

